Querying Integrated Scientific Observation and Measurement Data

### 2010-04-30 docs/algdraft/DataIntegration/DataIntegration.tex
Finish the rough writeup for query-rewriting.

\begin{itemize}
\item $Q_1$: Give me the datasets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the datasets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the datasets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the dataset that contains at least five distinct ``Picea rubens'' observations.
\item $Q_4$: Give me the datasets that have measurements with average ``area''
bigger than 5.0 square feet. 
\end{itemize}

@@

{\bf Experiement: show the error rate (e.g., precision, recall, $F_1$ of this method.}

@@

{\tt 
SELECT DISTINCT d.record\_id\\
FROM d\\
WHERE $f(d_{a_1} \cdots d_{a_m}) = true$\\
$[$GROUP BY$]$\\
$[$HAVING$]$
}

@@

{\bf TODO: Discuss how to deal with the distinct constraints. }
If there is distinct constraint, need to find the key measurements. 
then, perform ``group by'' operation.

{\bf TODO: Discuss how to deal with context}. 
If there is context, we need to use $A.ContextType$.

% Use the materialized database, we would have four tables for all the datasets: 
% entity, observation, measurement, context. 
% and 4 tables for entity types, observation types, measurement types,
% and context relationships.

\begin{example}\label{eg:query_reqerite_q1}
Take $Q_1$ as example. The {\tt concept} is ``observation'' and {\tt cond} is {\em species = ``Picea rubens''};

@@

The second step is to find the data file which really have this value. 
For this one, we find the data table, and do a selection on the table content. 
\end{example}

>>>>

\begin{itemize}
\item $Q_1$: Give me the datasets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the datasets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the datasets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the dataset that contains at least five distinct ``Picea rubens'' observations.
\item $Q_4$: Give me the datasets that have trees with average ``height''
  than 20in in year 2009. 
\item $Q_5$: an example on summarize one measure with respect to another.
This could take two forms: one where the context forms the groups
(like by year), and another where the two measurements are within the
same context (like finding correlations).
\end{itemize}

@@

{\bf TODO: An example to show that the naive method give wrong answers.}
{\bf Experiement: show the error rate (e.g., precision, recall, (or
  directly $F_1$) of this method.}

@@

[
{\tt 
SELECT DISTINCT d.record\_id\\
FROM d\\
WHERE $f(d_{a_1} \cdots d_{a_m}) = true$\\
$[$GROUP BY$]$\\
$[$HAVING$]$
}
]

@@

%Discuss how to deal with the distinct constraints.
Many times, a user may want to find {\em distinct} observations or
entities. In this case, the above simple solution cannot work
correctly because it treats observations in different rows as the same
observation. So, we need to use the annotation to figure out what
distinguishes one observation from another. 
When an observation type is denoted with {\em distinct yes} and some
of its measurement types are denoted with {\em key yes}, 
the values of the measurements on the key measurement types can
uniquely distinguish one observation from the other. 
To implement this, in SQL, we can perform ``group by'' operations on
the key measurement types of the observations.

%Discuss how to deal with context
The above description discusses cases without context. 
However, in real application, the observation has context. 
When the context is not {\em identifying}, i.e., the context value
does not affect the uniqueness of one observation, this can be
processed as the basic way. 
When the context is denoted with {\em identifying yes}, it means the
identity of one observation also depend on the contextual
observation. 
In this case, we need to figure out the all the key measurement types
of one observation, then the ``GROUP BY'' operation should be on all
these key measurement types.

{\bf TODO: put an example here.}

In this case, there is no need to change the SQL scripts.
the only change are the relevant attribute set $d_{a_1} \cdots
d_{a_m}$, which corresponds to not only the measurement types that
directly related to the observation, but also the measurement types
that are in the identifying context chain of the observation.

\begin{example}\label{eg:query_reqerite_q1}
{\bf TODO: refine this example} 
Take $Q_1$ as example. The {\tt concept} is ``observation'' and {\tt cond} is {\em species = ``Picea rubens''};

@@

The second step is to find the data file which really have this value. 
For this one, we find the data table, and do a selection on the table
content.

{\bf TODO: SQL here.} 
\end{example}
